---
title: "Visualization 3"
description: |
  This blog post shows the evolution of the third visual I created for this course
author:
  - name: Jenn Lewis
    url: https://15jennlee15.github.io/FinalProject_Website/
date: 03-16-2019
output:
  distill::distill_article:
    self_contained: false
categories:
     - Visualizations
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("flexdashboard")
#install.packages("tidyverse")
#install.packages("rio")
#install.packages("here")
#install.packages("rcartocolor")
#install.packages("devtools")
#install_github("datalorax/sundry")

library(tidyverse)
library(rio)
library(here)
library(rcartocolor)
library(devtools)
library(sundry)

data1 <- import(here("data", "Full_Data.csv")) 

data1$Group <- as.factor(data1$Group)
data1$Group <- recode(data1$Group, '1' = "Control Group", '2' = "BPD Group")

```


I created this visualization in RStudio using the following packages:  tidyverse, rio, sundry, here, devtools (to install sundry), and rcartocolor

## About the Data

See blog post "Visualization 1" for a description of the data!

## First Attempt at Visual 3

My initial attempt to make a scatterplot showing the relationship between a self-reported measure of emotion regulations (DERS) and inhibitory control as measured by the Stroop for both a BPD sample and healthy control sample.

```{r}
data1 %>%
  ggplot(aes(DERS, ColorWord_Score, color = Group)) +
  geom_point() +
  scale_colour_viridis_d() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(title = "The Relationship Between Emotion Regulation and Inhibitory Control",
       x = "Emotion Regulation Score (DERS)",
       y = "Inhibitory Control (Stroop Score)")
```

 As I examine this visual, I notice the outlier in the bottom corner, I also believe I can adjust some of the styling.  The colors specifically the yellow may be hard to see.
 
## Final Version of Visual 3

In this final version, I adjusted the colors to stand out more.  I also took some of the suggestions from my peer review and removed the outlier, and added a caption explaining what High DERS represents. I also made some stylistic changes.

This graph visualizes the relationship between inhibitory control and emotion regulation in women with and without BPD.  

```{r}
plot3 <- data1 %>%
  filter(Group != "NA" & ColorWord_Score > 20) %>%
  ggplot(aes(DERS, ColorWord_Score, color = Group)) +
  geom_point(alpha = 0.7) +
  scale_colour_carto_d() +
  geom_smooth(method = "lm", alpha = 0.2) +
  theme_minimal() +
  labs(title = "The Relationship Between Emotion Regulation and Inhibitory Control",
       x = "Emotion Regulation Score (DERS)",
       y = "Inhibitory Control (Stroop Score)",
       caption = "*Higher scores on the DERS represents more Emotion Regulation Difficulties")

plot3 + theme(plot.title = element_text(size = 12, hjust = 0.5))
```


